An Experimental Evaluation of Hybrid Vector Search (EA\&B)

### 2025-04-15 VLDB2026/main.tex

	Similar to the evaluation of \textit{Attribute Filtering} algorithms, we also evaluate 3 key metrics mentioned above in \textit{Range Filtering} algorithms.

@@

	In contrast, WinFilter has the longest construction time. This is mainly due to its tree-based structure, where each node requires an independent nearest neighbor graph. This leads to significant redundant computations and high resource costs, especially for large datasets.
	% 如表\ref{tab:Range Filtered}所示，在五种算法中，SeRF的构建时间最短，原因是其按属性值顺序增量构建的方式，仅需一次遍历并在局部图中添加边，结构简单、无冗余操作。而 WinFilter 的构建时间最长，主要由于其采用树状结构，在每个节点上需构建独立的近邻图索引，造成大量重复计算和资源开销，尤其在数据量大时更为明显。

	\textit{\textbf{Index size.}} Across the 3 datasets, iRange maintains a relatively small index size. It builds independent local neighbor graphs within each interval and merges indices dynamically during the query process. This prevents space explosion. On low-dimensional datasets (Deep, YT-Audio), SeRF compresses the index using sparse vector distributions, keeping it small. However, in the high-dimensional WIT dataset with high LID, vector connectivity increases. This reduces compression efficiency, leading to a sharp rise in index size. DSG maintains a stable index size across datasets but is generally larger. It uses edge structures with interval labels to support streaming insertions and range queries, increasing storage overhead.

	\textit{\textbf{Peak memory usage.}}
	In terms of peak memory usage, all 5 algorithms consume less memory on the first 2 datasets. However, memory usage increases significantly on WIT, likely due to its high dimensionality.

	SeRF achieves the lowest memory usage on low-dimensional datasets. Its Segment Graph efficiently compresses sparse vector spaces. However, on high-dimensional datasets with high LID, the “curse of dimensionality” causes distance concentration. As a result, the index structure of SeRF expands rapidly, reducing compression efficiency.

	In contrast, iRange achieves the lowest peak memory usage on WIT. It pre-builds graphs only for a limited number of intervals, avoiding explicit index construction for all possible query ranges. Its index structure remains stable in high dimensions without significant graph growth. This effectively controls index size, making iRange more advantageous for high-dimensional datasets.

@@

Except for DSG, most range filtering algorithms require the dataset to be pre-sorted before building the index. Moreover, these methods do not support multi-attribute range queries. Enhancing these methods to support an arbitrary number of range-filterable attributes remains an open research challenge.

>>>>

	Similar to the evaluation of \textit{Attribute-Filtering} algorithms, we also evaluate 3 key metrics mentioned above in \textit{Range Filtering} algorithms.

@@

	\textit{\textbf{Index construction time.}}
	Figure~\ref{fig:rangeFilter_build_time} shows the single-threaded index construction time. Faiss builds the index fastest.

	WinFilter and ACORN-\(\gamma\) take the longest.
	WinFilter relies on a tree structure, requiring separate nearest-neighbor graphs per node. This causes redundant computation and high costs, especially for large datasets.

	ACORN-\(\gamma\) expands the neighbor list of each node to provide more candidate paths, which increases both computation and storage costs. It also prunes distant neighbors to save storage space; however, this introduces additional overhead and slows down the construction process.

	\textit{\textbf{Index size.}} 
	As shown in Figure~\ref{fig:rangeFilter_index_size_mb}, Among all algorithms, Faiss has the smallest index structure because. 
	On low-dimensional datasets (Deep, YT-Audio), PASE requires the most storage. Its implementation likely wastes space due to alignment padding and inefficient encoding (e.g., Base64).

	SeRF maintains a compact index for low-dimensional data. However, for high-dimensional datasets (WIT), SeRF stores more edges and dynamic range data, significantly increasing its size.
	\textit{\textbf{Peak memory usage.}}
	As shown in Figure~\ref{fig:rangeFilter_memory_mb}, For low-dimensional datasets (deep, yt-audio), Faiss takes up the lowest memory due to: . 
	DSG and WinFilter take up significantly more memory than other algorithms due to.

	For high-dimensional datasets WIT, iRange achieves the lowest peak memory usage. It pre-builds graphs only for a limited number of intervals, avoiding explicit index construction for all possible query ranges. Its index structure remains stable in high dimensions without significant graph growth. This effectively controls index size, making iRange more advantageous for high-dimensional datasets.
	In contrast, unify takes up the highest memory due to

@@

Except for DSG, most range filtering algorithms require the dataset to be pre-sorted before building the index. Moreover, these methods do not support multi-attribute range queries. Enhancing these methods to support an arbitrary number of range filterable attributes remains an open research challenge.
